% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Della Vedova Gianluca},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Della Vedova Gianluca}
\author{}
\date{}

\begin{document}
\maketitle

\begin{quote}
Professore Associato Confermato (SSD INF/01 - Informatica)
\end{quote}

\begin{quote}
Dipartimento di Informatica, Sistemistica e Comunicazione • Università
degli Studi di Milano -- Bicocca
\end{quote}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\begin{description}
\tightlist
\item[10/2012-today]
Associate Professor, Dept. Informatics, Systems and Communications,
University of Milano -- Bicocca.
\item[10/2005-09/2012]
Associate Professor, School of Statistics, University of Milano --
Bicocca.
\item[05/2001-09/2005]
Assistant Professor, School of Statistics, University of Milano --
Bicocca.
\end{description}

\hypertarget{formazione}{%
\subsection{Formazione}\label{formazione}}

\begin{description}
\tightlist
\item[2001]
\textbf{Ph.D.~in Computer Science}; Università degli Studi di Milano
\item[1995]
\textbf{M.Sc. in Computer Science}; Università degli Studi di Milano.
\end{description}

\hypertarget{research}{%
\subsection{Research}\label{research}}

His research focuses on combinatorial algorithms in Bioinformatics, and
especially on sequence alignment, phylogeny inference, haplotyping, and
alternative splicing detection. He has also worked on graph
decomposition, clustering, and data privacy.

He has an MSc (1995) and a PhD (2001) in Computer Science from Univ. of
Milano.

He has published more than 30 journal papers with more than 1100
citations and h-index 15 (from Google Scholar). He has been the PI or
co-PI for several projects that have been funded more than 100.000€
overall. He is a member of the Editorial Board of ``Advances in
Bioinformatics'' journal, and he has been a guest editor of a special
issue of the Journal of Computer Science and Technology. He has been an
invited speaker at the Computaiblity in Europe 2017 conference.

H-index: 16. Citations: 1381
(\href{https://scholar.google.com/citations?user=0gaIFokAAAAJ\&hl=en\&oi=ao}{Google
Scholar}).

\hypertarget{selected-publications}{%
\subsection{Selected Publications}\label{selected-publications}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{4}
\tightlist
\item
  Ison, J, et al.~2015. ``Tools and data services registry: a community
  effort to document bioinformatics resources''. Nucleic acids research
  44 (D1), D38-D47.
\item
  Bonizzoni, Paola, Anna Paola Carrieri, Gianluca Della Vedova, and
  Gabriella Trucco. 2014. ``Explaining evolution via constrained
  persistent perfect phylogeny''. BMC Genomics 15 (6), S10.
\item
  Bonizzoni, Paola, Gianluca Della Vedova, and Riccardo Dondi. 2011.
  ``Anonymizing Binary and Small Tables Is Hard to Approximate.''
  Journal of Combinatorial Optimization 22 (1): 97--119.
\item
  Bonizzoni, Paola, Gianluca Della Vedova, Riccardo Dondi, and Jing Li.
  2003. ``The Haplotyping Problem: An Overview of Computational Models
  and Solutions.'' Journal of Computer Science and Technology 18 (6):
  675--88.
\item
  Bonizzoni, Paola, Gianluca Della Vedova, and Riccardo Dondi. 2005.
  ``Reconciling a Gene Tree to a Species Tree under the Duplication Cost
  Model.'' Theoretical Computer Science 347 (1-2): 36--53.
\item
  Borneman, J., M. Chrobak, Gianluca Della Vedova, A. Figueroa, and T
  Jiang. 2001. ``Probe Selection Algorithms with Applications in the
  Analysis of Microbial Communities.'' Bioinformatics 17 (Suppl 1):
  S39--48.
\item
  Bonizzoni, Paola, and Gianluca Della Vedova. 2001. ``The Complexity of
  Multiple Sequence Alignment with SP-Score That Is a Metric.''
  Theoretical Computer Science 259 (1-2): 63--79.
\end{enumerate}

\hypertarget{teaching}{%
\subsection{Teaching}\label{teaching}}

Since 2001, he has taught courses dedicated to students majoring in
Computer Science, Statistics, Bioinformatics, and Data Science.

\hypertarget{service}{%
\subsection{Service}\label{service}}

\begin{itemize}
\tightlist
\item
  E' stato membro del Editorial Board di ``Advances in Bioinformatics''.
\item
  Editor di special issue di ``Journal of Computer Science and
  Technology''
\item
  Membro di comitato di programma delle conferenze scientifiche CiE 2013
  2019 2020, SPIRE 2017, CPM 2019, WABI 2020.
\item
  Membro di comitato di programma dei workshop ``Workshop on Graph
  Assembly Algorithms for omics data'', ``Workshop on Algorithms in
  Molecular Biology''
\item
  Membro del Comitato direttivo del Centro di Produzione Multimediale di
  Ateneo
\item
  Supervisore di 3 tesi di dottorato e di oltre 10 tesi di laurea.
\item
  E' membro della Commissione Paritetica del Dipartimento di
  Informatica, Sistemistica e Comunicazione
\item
  E' Assicuratore di qualità del corso di laurea magistrale in Data
  Science.
\item
  E' stato delegato per l'e-learning del Preside della Facoltà di
  Scienze Statistiche.
\item
  E' stato rappresentante dell'Area informatica all'interno della
  Facoltà di Scienze Statistiche.
\item
  E' stato membro del comitato incaricato di progettare il corso di
  laurea triennale \emph{Statistica e Gestione delle Informazioni}.
\item
  E' stato il docente responsabile dei laboratori informatici della
  Facoltà di Scienze Statistiche.
\item
  E' stato responsabile del corso 167388 ``Laboratorio Complementare di
  Informatica per Statistici'', parte del progetto FSE 156165 ``Progetto
  Quadro Università degli Studi di Milano-Bicocca'', interamente
  finanziato da EU.
\item
  E' stato il responsabile per la Facoltà di Scienze Statistiche del
  corso ``Information Technology For Problem Solving (IT4PS)'',
  organizzato dalla Fondazione CRUI.
\end{itemize}

Milano, 17/02/2020

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\begin{quote}
\href{mailto:gianluca.dellavedova@unimib.it}{\nolinkurl{gianluca.dellavedova@unimib.it}}
• \url{https://gianluca.dellavedova.org}
\end{quote}

\begin{quote}
\href{https://orcid.org/0000-0001-5584-3089}{orcid 0000-0001-5584-3089}
\end{quote}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\begin{quote}
\href{mailto:gianluca.dellavedova@unimib.it}{\nolinkurl{gianluca.dellavedova@unimib.it}}
• \url{https://gianluca.dellavedova.org}
\end{quote}

\begin{quote}
\href{http://orcid.org/0000-0001-5584-3089}{orcid 0000-0001-5584-3089}
\end{quote}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\end{document}
